QECO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2024-11-04 Revision/Paper_Rev.tex
Rev_V8.1

\begin{itemize}
	\item \textit{Task Offloading Problem in the MEC System:} \textcolor{red}{ We consider queuing systems for MDs and ENs and formulate the task offloading problem as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). We aim to maximize the long-term QoE, which reflects the task completion rate, task delay, and energy consumption of MDs. Our problem formulation effectively utilizes the resources and properly handles the dynamic fluctuations of workload at ENs. }

	%We formulate the task offloading problem in queuing-based MEC as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). Given task completion, task delay, and energy consumption as the QoE requirements of MDs, this approach concentrates on maximizing the long-term QoE, by effectively utilizing the resources and properly handling the dynamic nature of workload at ENs.

@@

\end{itemize}

The structure of this paper is as follows.  \textcolor{red}{Section~\ref{section:II} reviews the related works.} Section~\ref{section:III} presents the system model, followed by the problem formulation in Section ~\ref{section:IV}. In Section~\ref{section:V}, we present the algorithm, while Section~\ref{section:VI} provides an evaluation of its performance. Finally, we conclude in Section~\ref{section:VII}.

\newpage.\newpage

\section{Related Work}
\label{section:II}
%\textcolor{black}{We explore state-of-the-art techniques characteristic to discuss how they address task offloading problems in scenarios that need to process massive data with limited devices in MEC, such as the Internet of Things (IoT), Internet of Vehicles (IoV), and Industrial IoT (IIoT)  environment. 
%Internet of Things (IoT) \cite{dai2020edge}, \cite{liao2023online}, \cite{Bolourian-WCL24}, 
%Internet of Vehicles (IoV) \cite{Bolourian-WCL24}, 
%and Industrial IoT (IIoT)  environment \cite{gong2022edge}, \cite{wu2023multi}, \cite{wu2024combining}
%, while uncertain requirements and sensitive latency properties of computation tasks pose nontrivial challenges in MEC systems with limited resources. 
% \textcolor{blue}{In recent years, edge computing has become a popular method for scenarios that need to process massive data with weak devices, such as IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{lin2022multi}, \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive} environment. We explore state-of-the-art RL-based techniques characteristic to discuss how they address task offloading problems in MEC. Table~\ref{table1} provides an intuitive comparison of the related works.}
  In recent years, numerous studies have focused on computation offloading in MEC, covering a broad range of applications, including IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{lin2022multi}, \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive}. 
    %To take advantage of RL methods for computation offloading in MEC, most studies transfer the problem to MDP to model the interactions of the environment. 
    To effectively tackle the challenges of MEC arising from the ever-changing nature of networks, recent research highlights the effectiveness of DRL in adapting to environmental changes and learning optimal strategies. In this section, we explore the current DRL-based state-of-the-art works and discuss their strengths and limitations. Table~\ref{table1} provides an intuitive comparison of the these works.
    \subsection{DRL-based Optimization Methods for MEC}

Mainly, there are two primary categories of related researchs based on the target objectives in resource management for MEC, single-objective optimization and multi-objective optimization methods, where DRL are mainstream optimization techniques. 
\subsubsection{Single-Objective Optimization} 
There has been a large amount of research studying single-objective optimization problems in the context of MEC, where only one objective is considered for optimization, e.g., delay or enery consumption.
%\textbf{Number of compeleted tasks optimization.}
%\textbf{Delay-optimal algorithms.}  
Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources. due to the dynamic nature of MEC systems, Li et al. [25] proposed an integrated optimization learning framework called DROO, which leverages past experience and model informa- tion to provide fast and robust real-time offloading control.
Li et al. [39] proposed an offloading strategy through the DRL algorithm to design the system, where a DQN is used to auto- matically learn offloading decisions, optimize system performance, and train a neural network (NN) to predict offloading actions. 
To reduce the total delay and waiting time of mobile vehicle tasks, Tang et al. [40] established a dynamic offloading model of mul- tiple mobile vehicles, dividing its tasks into sequential subtasks, and making the offloading decision more refined.

%\textbf{Energy-efficient algorithms.} 
To minimize energy consumption, Munir \textit{et al.} \cite{munir2021multi} developed a semi-distributed approach using a multi-agent RL framework for self-powered MEC. 
Zhou \textit{et al.} in \cite{zhou2021deep} proposed a Q-learning approach, which is an extension of RL to achieve optimal resource allocation strategies and computation offloading.
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to concurrently optimize resource allocation and computation offloading.
	Wang et al. [52] investigated the computation offload- ing problem in MEC systems with mobile awareness in cellular networks, proposing a joint task offloading and computation alocation based on mobile-aware migration. Their approach focuses on the task offloading cost caused by user mobility.	In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation.

@@

To reflecting real-world scenarios more accurately, the authors applied queuing theory to model MEC systems. It helps in modeling the dynamic nature of network, allowing MEC systems to allocate computational and transmission resources more efficiently.
 	Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To address the privacy-aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} modeled the problem as an MDP and proposed a DQN-based method to optimize the computation rate and energy consumption in a queuing-based IIoT network.
	 Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. 
  	 	 Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the IIoT systems to jointly optimize task offloading and resource allocation to achieve lower energy consumption and decreased task delay.
 To optimize delay and energy consumption, Wu \textit{et al.} in \cite{wu2023multi} investigated the computation offloading problem in a queuing-based MEC IIoT system. They modeled the problem as a POMDP and proposed a multi-agent proximal policy optimization (PPO)-based method to obtain the optimal offloading strategy in dynamic environments.
 Tang \textit{et al.} in \cite{9253665} investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. 
 In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading.

 \textcolor{black}{\subsection{Analysis and Motivation}}

@@

  \subsubsection{System modeling}
  Some existing works, see \cite{huang2019deep}--\cite{zhou2021deep}, primarily investigate simplified MEC networks, focusing on single MEC network models, considering only single-server or single-user scenarios, which may not always reflect real-world scenarios. Some other in \cite{huang2019deep}, \cite{liu2021learn} do not account for delay-sensitive tasks and may overlook critical performance factors in real-time applications. Comparing with the aforementioned works [10]–[15], we consider a different and more realistic MEC scenario involving multuple MDs and ENs, which take into account delay-sensitive tasks with processing deadlines. This is challenging to address, because the processing deadlines will affect the load level dynamics at the edge nodes and hence affect the delay of the offloaded tasks.
Besides, some works with more advanced scarios do not take into account the underlying resources schetuling systems. As a result, the processing and transmission of each task should always be accomplished within one time slot (or before the arrival of the next task), which may not always be guaranteed in practice. Different from those works [10]–[15], we consider delay-sensitive tasks together with queuing systems and take into account the practical scenario where the processing and transmission of each task can continue for multiple time slots.

@@

  %In more advanced scenarios, some studies explored task offloading problems

  % \textbf{Dynamic network environment considering.}

@@

 \subsubsection{Algorithme design}
 On the other hand, due to the limited processing capacities of ENs, the processing capacity that an EN allocates to an MD depends on the EN's workload each time. Treating this dynamic workload as part of the environment can negatively impact system performance and requires careful consideration. Some studies did not take into account the challenges of a dynamic network environment. For instance, in \cite{chen2021drl}, MDs are assumed to have sufficient transmission and computation resources from ENs, which may not always align with real-world conditions. While the works of \cite{wu2024privacy}--\cite{zhao2019deep} investigated the limited resources in MEC, they did not consider the dynamic workload at ENs in decision making processes, imposing significant challenges in overall system performance. Although some existing works \cite{gong2022edge}–\cite{gao2022large} have addressed the load levels at the edge nodes and proposed task offloading algorithms, however, these algorithms in require global information of the system and may incur high signaling overhead. Different from those works [17]–[20], we aim to propose a DRL-based distributed algorithm that can address the unknown load dynamics at the edge nodes. It should also enable each MDs to make its offloading decision without knowing the information (e.g., task models, offloading decisions) of other mobile devices.

@@

 \newpage
 These works in \cite{huang2019deep}, \cite{liu2021learn} do not account for delay-sensitive tasks and may overlook critical performance factors in real-time applications. Besides, these studies \cite{huang2019deep}--\cite{zhou2021deep} primarily investigate simplified MEC networks, focusing on single MEC network models, considering only single-server or single-user scenarios, which may not always reflect real-world scenarios. 
 Moreover, due to the absence of robust underlying resource scheduling systems, issues such as data congestion during transmission and resource allocation can arise, leading to suboptimal system performance. Therefore, more accurate MEC models are needed to address the complexities of dynamic resource scheduling in large-scale scenarios.
 These works did not take into account the challenges of a dynamic network environment. For instance, in \cite{chen2021drl}, MDs are assumed to have sufficient transmission and computation resources from ENs, which may not always align with real-world conditions. While the works of \cite{wu2024privacy}--\cite{zhao2019deep} investigated the limited resources in MEC, they did not consider the dynamic workload at ENs in decision making processes, imposing significant challenges in overall system performance.
 	 In above research \cite{munir2021multi}--\cite{huang2021deadline}, while  models were employed in several parts of the system, the offloading process was not approached as a holistic queuing system. 
 	 	 Due to the limited processing capacities of ENs, the processing capacity that an EN allocates to an MD depends on the EN's workload each time. Treating this dynamic workload as part of the environment can negatively impact system performance and requires careful consideration. Specifically, offloaded tasks may experience significant processing delays or be dropped if their deadlines expire. Some existing works have addressed the load levels at the edge nodes and proposed task offloading algorithms.	 
 	 	 However, these algorithms in \cite{gong2022edge}–\cite{gao2022large} require global information of the system and may incur high signaling overhead.

@@

\begin{table*}[htbp]\textcolor{blue}{
		\renewcommand{\arraystretch}{0.7}
		\captionsetup{name=TABLE}
		\caption{Comparison of Related Works}
		\scalebox{1}{%
			\begin{tabular}{ lp{3cm}p{5.3cm}p{2cm}p{4.8cm}l} 
				\toprule
				\textbf{Paper}  &  \textbf{Problem} &  \textbf{Objective} & \textbf{Method} &  \textbf{Limitation}  \\ \midrule
				\cite{liu2021learn}  & Resource allocation and computation offloading & Minimize execution cost  & Multi-agent actor-critic network &  Not consider delay-sensitive tasks \\\midrule
				\cite{wu2023computation}   & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & Multi-agent RL & Not consider tasks with maximum delay tolerance\\\midrule
				\cite{zhou2021deep} & Resource allocation and computation offloading & Minimize energy consumption of the entire system & DDQN & Only for a single MEC system model \\\midrule				
				%\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
				\cite{wu2024privacy} & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service.  & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{dai2020edge} & Resource allocation and computation offloading & Minimize energy consumption & DDPG & Not take user's demand into consideration \\\midrule
				\cite{zhao2019deep} & User association and resource allocation & Optimize downlink utility  & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
				\cite{liao2023online}  & Online computation offloading & Minimize computation delay and energy consumption & DDQN & Only for a single MEC system model  \\\midrule
				\cite{wu2024combining}   & Privacy aware computation offloading & Optimization of computation rate and energy consumption  & Actor-critic network  & Not take user's demand into consideration \\\midrule
				\cite{Bolourian-WCL24}  & Energy harvesting  computation offloading & Minimize drop rate and energy consumption.  & DQN & Not consider computation delay  \\\midrule
				\cite{huang2021deadline}  & Computation offloading & Optimize drop rate and energy consumption & DDPG & Not consider computation delay \\ \midrule
				\cite{gong2022edge} & Resource allocation and computation offloading & Minimize task computation delay and energy consumption  & DRL & Not take user's demand into consideration\\\midrule
				\cite{sun2024hierarchical} & Computation offloading and service caching & Minimize average service delay & Hierarchical DRL & Not consider system energy consumtion \\\midrule
				\cite{wu2023multi}  & Computation offloading & Optimize delay and energy consumption  & Multi-agent PPO&  Not consider tasks with maximum delay tolerance \\\midrule
				\cite{9253665}  &  Computation offloading & Minimize drop rate and computation delay.  & D3QN + LSTM & Not consider system energy consumtion\\\midrule
				\cite{gao2022large}  & Decentralized computation offloading & Optimize system cost and completion rates  & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				QECO  &  Computation offloading & Maximize the QoE of each MD individually & D3QN + LSTM & \\
				\toprule
		\end{tabular}}
		\label{table1}}
\end{table*}

\newpage.\newpage

@@

The EN randomly selects a sample collection of experiences from the replay buffer, denoted as $\mathcal{N}$. For each experience $n \in \mathcal{N}$, it calculates the value of $	\hat{Q}^{\text{T}}_{i,n}$. This value represents the QoE in experience $n$ and includes a discounted Q-value of the action anticipated to be taken in the subsequent state of experience $n$, according to the network $\textit{Net}^\text{T}_i$, given by
\begin{alignat}{1}
	\hat{Q}_{i,n}^{\text{T}} = \boldsymbol{q}_i(n) + \gamma Q_i^{\text{T}}(\boldsymbol{s}_i(n+1)), \tilde{\boldsymbol{a}}_n; \theta_i^{\text{T}}),
	\label{29}  
\end{alignat}  
where $\tilde{\boldsymbol{a}}_n$ denotes the optimal action for the state $\boldsymbol{s}_i(n+1)$ based on its highest Q-value under $\textit{Net}_i^{\text{E}}$, as given by:
\begin{alignat}{1}
	\tilde{\boldsymbol{a}}_n = \text{arg} \; \max_{\boldsymbol{a} \in \mathcal{A}} \; Q_i^{\text{E}}(\boldsymbol{s}_i(n+1), \boldsymbol{a}; \theta_i^{\text{E}}).
	\label{30}  
\end{alignat} 
In particular, regarding experience $n$, the target-Q value $\hat{Q}_{i,n}^{\text{T}}$ represents the long-term QoE for action $\boldsymbol{a}_i(n)$ under state $\boldsymbol{s}_i(n)$. This value corresponds to the QoE observed in experience $n$, as well as the approximate expected upcoming QoE. %based on $\textit{Net}_i^{\text{T}}_i$, i.e., $\gamma Q_i^{\text{T}}(\mathcal{S}_i(i+1)), \mathcal{A}_i^{\text{N}}; \theta_i^{\text{T}})$
%Based on the set $\mathcal{N}$, the EN trains the MD's neural network using previous sample experiences. Simultaneously, it updates $\theta^{\text{E}}_i$ in $\textit{Net}_i^{\text{E}}$ and computes vector $\hat{\mathbf{Q}}_i^{\text{T}} = (\hat{Q}^{\text{T}}_{i,n})_{n \in \mathcal{N}}$. The key idea of updating $\textit{Net}_i^{\text{E}}$ is to minimize the disparity in Q-values between $\textit{Net}_i^{\text{E}}$ and $\textit{Net}_i^{\text{T}}$, as indicated by the following loss function:
\textcolor{red}{Based on the set $\mathcal{N}$, the EN computes the vector $\hat{\mathbf{Q}}_i^{\text{T}} = (\hat{Q}^{\text{T}}_{i,n})_{n \in \mathcal{N}}$ and trains the MD's neural network (steps 11-21 of Algorithm~\ref{alg:cap2}) to update the parameter vector $\theta^{\text{E}}_i$ in $\textit{Net}_i^{\text{E}}$ for the next MD's \textit{UpdateRequest}.} The key idea of updating $\textit{Net}_i^{\text{E}}$ is to minimize the disparity in Q-values between $\textit{Net}_i^{\text{E}}$ and $\textit{Net}_i^{\text{T}}$, as indicated by the following loss function:

>>>>

\begin{itemize}
	\item \textit{Task Offloading Problem in the MEC System:} \textcolor{red}{ We consider queuing systems for MDs and ENs and formulate the task offloading problem as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). We aim to maximize the long-term QoE, which reflects the task completion rate, task delay, and energy consumption of MDs. Our problem formulation effectively utilizes the resources and properly handles the dynamic fluctuations of workload at ENs. }
	%We formulate the task offloading problem in queuing-based MEC as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). Given task completion, task delay, and energy consumption as the QoE requirements of MDs, this approach concentrates on maximizing the long-term QoE, by effectively utilizing the resources and properly handling the dynamic nature of workload at ENs.

@@

\end{itemize}
The structure of this paper is as follows.  \textcolor{red}{Section~\ref{section:II} reviews the related works.} Section~\ref{section:III} presents the system model, followed by the problem formulation in Section ~\ref{section:IV}. In Section~\ref{section:V}, we present the algorithm, while Section~\ref{section:VI} provides an evaluation of its performance. Finally, we conclude in Section~\ref{section:VII}.
\section{Related Work}
\label{section:II}
%\textcolor{black}{We explore state-of-the-art techniques characteristic to discuss how they address task offloading problems in scenarios that need to process massive data with limited devices in MEC, such as the Internet of Things (IoT), Internet of Vehicles (IoV), and Industrial IoT (IIoT)  environment. 
%Internet of Things (IoT) \cite{dai2020edge}, \cite{liao2023online}, \cite{Bolourian-WCL24}, 
%Internet of Vehicles (IoV) \cite{Bolourian-WCL24}, 
%and Industrial IoT (IIoT)  environment \cite{gong2022edge}, \cite{wu2023multi}, \cite{wu2024combining}
%, while uncertain requirements and sensitive latency properties of computation tasks pose nontrivial challenges in MEC systems with limited resources. 
% \textcolor{blue}{In recent years, edge computing has become a popular method for scenarios that need to process massive data with weak devices, such as IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{lin2022multi}, \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive} environment. We explore state-of-the-art RL-based techniques characteristic to discuss how they address task offloading problems in MEC. Table~\ref{table1} provides an intuitive comparison of the related works.}
  In recent years, numerous studies have focused on computation offloading in MEC, covering a broad range of applications, including IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{lin2022multi}, \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive}. 
    %To take advantage of RL methods for computation offloading in MEC, most studies transfer the problem to MDP to model the interactions of the environment. 
    To effectively tackle the challenges of MEC arising from the ever-changing nature of networks, recent research highlights the effectiveness of DRL in adapting to environmental changes and learning optimal strategies. In this section, we explore the current DRL-based state-of-the-art works and discuss their strengths and limitations. Table~\ref{table1} provides an intuitive comparison of the these works.
    \subsection{DRL-based Optimization Methods for MEC}

Mainly, there are two primary categories of related researchs based on the target objectives in resource management for MEC, single-objective optimization and multi-objective optimization methods, where DRL are mainstream optimization techniques. 
\subsubsection{Single-Objective Optimization} 
There has been a large amount of research studying single-objective optimization problems in the context of MEC, where only one objective is considered for optimization, e.g., delay or enery consumption.
%\textbf{Number of compeleted tasks optimization.}
%\textbf{Delay-optimal algorithms.}  
Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources. Due to the dynamic nature of MEC systems, Li \textit{et al.} in \cite{li2022integrated} proposed an integrated optimization learning framework called DROO, which leverages past experience and model informa- tion to provide fast and robust real-time offloading control.
In \cite{li2020joint} the authors proposed an offloading strategy through the DRL algorithm to design the system, where a DQN is used to automatically learn offloading decisions, optimize system performance, and train a neural network (NN) to predict offloading actions. 
To reduce the total delay and waiting time of mobile vehicle tasks, Tang  \textit{et al.} in \cite{tang2022double}  established a dynamic offloading model of mul- tiple mobile vehicles, dividing its tasks into sequential subtasks, and making the offloading decision more refined.

%\textbf{Energy-efficient algorithms.} 
To minimize energy consumption, Munir \textit{et al.} \cite{munir2021multi} developed a semi-distributed approach using a multi-agent RL framework for self-powered MEC. 
Zhou \textit{et al.} in \cite{zhou2021deep} proposed a Q-learning approach, which is an extension of RL to achieve optimal resource allocation strategies and computation offloading.
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to concurrently optimize resource allocation and computation offloading.
	Wang \textit{et al.} in \cite{wang2020reinforcement} investigated the computation offload- ing problem in MEC systems with mobile awareness in cellular networks, proposing a joint task offloading and computation alocation based on mobile-aware migration. Their approach focuses on the task offloading cost caused by user mobility.	In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation.

	 \begin{table*}[htbp]\textcolor{blue}{
			\renewcommand{\arraystretch}{0.2}
			\captionsetup{name=TABLE}
			\caption{Comparison of Related Works}
			\scalebox{1}{%
				\begin{tabular}{ lp{4cm}p{4.6cm}p{2cm}p{4.7cm}l} 
					\toprule
					\textbf{Paper}  &  \textbf{Problem} &  \textbf{Objective} & \textbf{Method} &  \textbf{Limitation}  \\ \midrule
					\cite{li2022integrated} & Online computation offloading & Minimize computation delay & Actor-critic network & Not consider system energy consumtion \\ \midrule
					\cite{li2020joint} & Caching and computation offloading & Minimize delay of the whole system &  Multi-agent DQN & Not consider system energy consumtion \\ \midrule
					\cite{tang2022double}  & Computation offloading & Minimize total delay & DDQN & Not consider system energy consumtion \\ \midrule
					\cite{wang2020reinforcement} & Mobility aware computation offloading & Minimize energy consumtion & Q-learning & Not consider computation delay\\ \midrule
					\cite{liu2021learn}  & Resource allocation and computation offloading & Minimize execution cost  & Multi-agent actor-critic network &  Not consider delay-sensitive tasks \\\midrule
					\cite{wu2023computation}   & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & Multi-agent RL & Not consider tasks with maximum delay tolerance\\\midrule
					\cite{zhou2021deep} & Resource allocation and computation offloading & Minimize energy consumption of the entire system & DDQN & Only for a single MEC system model \\\midrule				
					%\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
					\cite{wu2024privacy} & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service.  & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
					\cite{dai2020edge} & Resource allocation and computation offloading & Minimize energy consumption & DDPG & Not take user's demand into consideration \\\midrule
					\cite{zhao2019deep} & User association and resource allocation & Optimize downlink utility  & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
					\cite{liao2023online}  & Online computation offloading & Minimize computation delay and energy consumption & DDQN & Only for a single MEC system model  \\\midrule
					\cite{wu2024combining}   & Privacy aware computation offloading & Optimization of computation rate and energy consumption  & Actor-critic network  & Not take user's demand into consideration \\\midrule
					\cite{Bolourian-WCL24}  & Energy harvesting  computation offloading & Minimize drop rate and energy consumption.  & DQN & Not consider computation delay  \\\midrule
					\cite{huang2021deadline}  & Computation offloading & Optimize drop rate and energy consumption & DDPG & Not consider computation delay \\ \midrule
					\cite{gong2022edge} & Resource allocation and computation offloading & Minimize task computation delay and energy consumption  & DRL & Not take user's demand into consideration\\\midrule
					\cite{sun2024hierarchical} & Computation offloading and service caching & Minimize average service delay & Hierarchical DRL & Not consider system energy consumtion \\\midrule
					\cite{wu2023multi}  & Computation offloading & Optimize delay and energy consumption  & Multi-agent PPO&  Not consider tasks with maximum delay tolerance \\\midrule
					\cite{9253665}  &  Computation offloading & Minimize drop rate and computation delay.  & D3QN + LSTM & Not consider system energy consumtion\\\midrule
					\cite{gao2022large}  & Decentralized computation offloading & Optimize system cost and completion rates  & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
					QECO  &  Computation offloading & Maximize the QoE of each MD individually & D3QN + LSTM & \\
					\toprule
			\end{tabular}}
			\label{table1}}
	\end{table*}

@@

To reflecting real-world scenarios more accurately, the authors applied queuing theory to model MEC systems. It helps in modeling the dynamic nature of network, allowing MEC systems to allocate computational and transmission resources more efficiently.
 	Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To address the privacy-aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} modeled the problem as an MDP and proposed a DQN-based method to optimize the computation rate and energy consumption in a queuing-based IIoT network.
	 Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. 
  	 	 Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the IIoT systems to jointly optimize task offloading and resource allocation to achieve lower energy consumption and decreased task delay.
 To optimize delay and energy consumption, Wu \textit{et al.} in \cite{wu2023multi} investigated the computation offloading problem in a queuing-based MEC IIoT system. They modeled the problem as a POMDP and proposed a multi-agent proximal policy optimization (PPO)-based method to obtain the optimal offloading strategy in dynamic environments.
 Tang \textit{et al.} in \cite{9253665} investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. 
 In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading.
 \textcolor{black}{\subsection{Analysis and Motivation}}

@@

  \subsubsection{System modeling}
  Some existing works, see \cite{huang2019deep}--\cite{zhou2021deep}, primarily investigate simplified MEC networks, focusing on single MEC network models, considering only single-server or single-user scenarios, which may not always reflect real-world scenarios. Some other in \cite{huang2019deep}, \cite{liu2021learn} do not account for delay-sensitive tasks and may overlook critical performance factors in real-time applications. Comparing with the aforementioned works [10]–[15], we consider a different and more realistic MEC scenario involving multuple MDs and ENs, which take into account delay-sensitive tasks with processing deadlines. This is challenging to address, because the processing deadlines will affect the load level dynamics at the edge nodes and hence affect the delay of the offloaded tasks.
Besides, some works with more advanced scarios do not take into account the underlying resources schetuling systems. As a result, the processing and transmission of each task should always be accomplished within one time slot (or before the arrival of the next task), which may not always be guaranteed in practice. Different from those works [10]–[15], we consider delay-sensitive tasks together with queuing systems and take into account the practical scenario where the processing and transmission of each task can continue for multiple time slots.

@@

@@

 \subsubsection{Algorithme design}
 On the other hand, due to the limited processing capacities of ENs, the processing capacity that an EN allocates to an MD depends on the EN's workload each time. Treating this dynamic workload as part of the environment can negatively impact system performance and requires careful consideration. Some studies did not take into account the challenges of a dynamic network environment. For instance, in \cite{chen2021drl}, MDs are assumed to have sufficient transmission and computation resources from ENs, which may not always align with real-world conditions. While the works of \cite{wu2024privacy}--\cite{zhao2019deep} investigated the limited resources in MEC, they did not consider the dynamic workload at ENs in decision making processes, imposing significant challenges in overall system performance. Although some existing works \cite{gong2022edge}–\cite{gao2022large} have addressed the load levels at the edge nodes and proposed task offloading algorithms, however, these algorithms in require global information of the system and may incur high signaling overhead. Different from those works [17]–[20], we aim to propose a DRL-based distributed algorithm that can address the unknown load dynamics at the edge nodes. It should also enable each MDs to make its offloading decision without knowing the information (e.g., task models, offloading decisions) of other MDs.

@@

@@

@@

The EN randomly selects a sample collection of experiences from the replay buffer, denoted as $\mathcal{N}$. For each experience $n \in \mathcal{N}$, it calculates the value of $	\hat{Q}^{\text{T}}_{i,n}$. This value represents the QoE in experience $n$ and includes a discounted Q-value of the action anticipated to be taken in the subsequent state of experience $n$, according to the network $\textit{Net}^\text{T}_i$, given by
\begin{alignat}{1}
	\hat{Q}_{i,n}^{\text{T}} = \boldsymbol{q}_i(n) + \gamma Q_i^{\text{T}}(\boldsymbol{s}_i(n+1)), \tilde{\boldsymbol{a}}_n; \theta_i^{\text{T}}),
	\label{29}  
\end{alignat}  
where $\tilde{\boldsymbol{a}}_n$ denotes the optimal action for the state $\boldsymbol{s}_i(n+1)$ based on its highest Q-value under $\textit{Net}_i^{\text{E}}$, as given by:
\begin{alignat}{1}
	\tilde{\boldsymbol{a}}_n = \text{arg} \; \max_{\boldsymbol{a} \in \mathcal{A}} \; Q_i^{\text{E}}(\boldsymbol{s}_i(n+1), \boldsymbol{a}; \theta_i^{\text{E}}).
	\label{30}  
\end{alignat} 
In particular, regarding experience $n$, the target-Q value $\hat{Q}_{i,n}^{\text{T}}$ represents the long-term QoE for action $\boldsymbol{a}_i(n)$ under state $\boldsymbol{s}_i(n)$. This value corresponds to the QoE observed in experience $n$, as well as the approximate expected upcoming QoE. %based on $\textit{Net}_i^{\text{T}}_i$, i.e., $\gamma Q_i^{\text{T}}(\mathcal{S}_i(i+1)), \mathcal{A}_i^{\text{N}}; \theta_i^{\text{T}})$
%Based on the set $\mathcal{N}$, the EN trains the MD's neural network using previous sample experiences. Simultaneously, it updates $\theta^{\text{E}}_i$ in $\textit{Net}_i^{\text{E}}$ and computes vector $\hat{\mathbf{Q}}_i^{\text{T}} = (\hat{Q}^{\text{T}}_{i,n})_{n \in \mathcal{N}}$. The key idea of updating $\textit{Net}_i^{\text{E}}$ is to minimize the disparity in Q-values between $\textit{Net}_i^{\text{E}}$ and $\textit{Net}_i^{\text{T}}$, as indicated by the following loss function:
\textcolor{red}{Based on the set $\mathcal{N}$, the EN computes the vector $\hat{\mathbf{Q}}_i^{\text{T}} = (\hat{Q}^{\text{T}}_{i,n})_{n \in \mathcal{N}}$ and trains the MD's neural network (steps 11-21 of Algorithm~\ref{alg:cap2}) to update the parameter vector $\theta^{\text{E}}_i$ in $\textit{Net}_i^{\text{E}}$ for the next MD's \textit{UpdateRequest}.} The key idea of updating $\textit{Net}_i^{\text{E}}$ is to minimize the disparity in Q-values between $\textit{Net}_i^{\text{E}}$ and $\textit{Net}_i^{\text{T}}$, as indicated by the following loss function:
\begin{alignat}{1}
	\hspace{-2mm}L(\theta_i^{\text{E}},\hat{\mathbf{Q}}_i^{\text{T}}) = {1 \over |\mathcal{N}| }\hspace{-1mm} \sum\limits_{n \in \mathcal{N}} \bigg(Q_i^{\text{E}}(\boldsymbol{s}_i(n) , \boldsymbol{a}_i(n); \theta_i^{\text{E}} ) -   \hat{Q}^{\text{T}}_{i,n}  \bigg)^2.
	\label{31}  
\end{alignat}  
In every \textit{ReplaceThreshold} iterations, the update of $\textit{Net}_i^{\text{T}}$ will involve duplicating the parameters from $\textit{Net}_i^{\text{E}}$ ($\theta_i^{T} = \theta_i^{E}$). The objective is to consistently update the network parameter $\theta_i^{T}$ in $\textit{Net}_i^{\text{T}}$, which enhances the approximation of the long-term QoE when computing the target Q-values in \eqref{28}.\\
